---
title: "Characteristics of Recent Mexican Immigrants to the United States that Influence Household Income"
subtitle: "Regression Analysis"
author: "Ben 10"
date: "November 20, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-packages}
library(tidyverse)
library(broom)
library(leaps)
library(knitr)
library(dplyr)
library(rms)
```

```{r data}
data <- read_csv("/cloud/project/03-regression-analysis/data.csv")
```

## 1. Introduction

### 1.1 Objective

We are aiming to build a model to determine which characteristics of Mexican immmigrants to the United States, specifically California, well-explain variation in household income. 

### 1.2 Description of Dataset

The dataset is from The Mexican Migration Project (MMP, *see References below for confidentiality terms). It was created in 1982 by an interdisciplinary team of researchers to further our understanding of the complex process of Mexican migration to the United States. The project is a binational research effort co-directed by Jorge Durand, professor of Social Anthropology at the University of Guadalajara (Mexico), and Douglas S. Massey, professor of Sociology and Public Affairs, with a joint appointment in the Woodrow Wilson School, at Princeton University (US).

Since its inception, the MMP's main focus has been to gather social as well as economic information on Mexican-US migration. The data collected has been compiled in a comprehensive database that is available to the public free of charge for research and educational purposes through its web-site. The MMP uses the ethnosurvey approach to gather data: in winter months, they randomly sample households in communities throughout Mexico, surveying household heads and members about their first and last trip to the US, as well as economic and demographic information. They then conduct the same survey in destination areas in the US, sampling migrants from the same communities they survey in Mexico but who have not returned to Mexico. Thus, the sample of migrants includes residents in both Mexico and the US.

The MMP170 Database contains an initial file with general demographic, economic, and migratory information for each member of a surveyed household (PERS). Pers170 has 132 variables and 176701 observations,and hence it is very large. Therefore, we selected 17 meaningful variables and filtered out rows that contain N/A's to create a new dataset labeled `data`.

### 1.3 Method

We will build a multiple linear regression model to predict household income considering the following variables: `sex`, `relhead`, `age`, `statebrn`, `marstat`, `edyrs`, `occtype`, `usdur1`, `usdurl`, `usdoc1`, `uscity`, `yrborn`. 

"X1": Number of observation

"sex": Sex
  
"relhead": Relationship to household head

"yrborn": Year of birth

"age": Age

"statebrn": State of birth

"marstat": Marital status

"edyrs": School years completed

"occ": Principal occupation

"hhincome" : Household income

"usstate1": First US migration: State of residence

"usstatel": Latest US migration: State of residence

"usplace1": First US migration: City of residence (in place codes)

"usplacel": Latest US migration: City of residence (in place codes)

"usdur1": First US migration: Duration (in months)

"usdurl": Latest US migration: Duration (in months)

"usdoc1": Type of documentation 

"occtype": Category of occupation

"uscity": City of residence during first US migration

Our response variable is household income: the total income for a single household, reported in $USD. We chose to use the multiple linear regression because our response variable is numeric, and there are multiple predictor variables.


## 2. Exploratory Data Analysis

### 2.1 Data Cleaning

Due to the complexity of our original data, we did not include data cleaning in the analysis. For more information, please see our proposal, where all the data cleaning happens.

However, we did make some adjustment according to the feedback that there is large imbalance of the amount of data between regions, and that the distribution of the response variable is not normal. Below is the update on our data cleaning:

### 2.2 Updated Data Exploration

#### 2.2.1 Filter Only Immigrants in California

Accoridng to our previous data exploration, we found that the overwhelming majority of immigrants settled in California, as shown in the graph below:

```{r}
ggplot(data = data, aes(x = uscity)) + 
  geom_bar() +
  coord_flip()+
  labs(title = "Distribution of City of Residence", x= "City of Residence", y = "Count")
```

Hence, we decided to concentrate on California alone. Since the originial dataset is large, we have enough data left in California alone to produce meaningful analysis.

```{r caldata}
caldata <- data %>%
  filter(usstatel == "California")

caldata <- caldata%>%
  na.omit()

```

#### 2.2.2 Cut Household Income Groups

Originally, the distribution of log(Household Income)- our response variable- was bimodal and had a median of 412,647 dollars. It almost looks like 3 separate distributions:

```{r nonfilt-income-plot}
ggplot(data = caldata, aes(x = log(hhincome))) +
  geom_histogram(binwidth = .7)+
  labs(title = "Distribution of Log(Household Income)", x= "log(Household Income)", y = "Count")
```

We determined that 412,647 dollars is an absurdly high median income for a survey of largely undocumented immigrants in the US and believe that a significant chunk of the high incomes were actually recorded in pesos. However, since the data is taken across a period of 10 years, the exchange rate between pesos and USD had increased significantly. Hence, we cannot simply convert all the data that are larger into USD. 

Therefore, we decided to filter out the incomes above 60,000 to remove what appears to be a second distribution of incomes in pesos. We will also remove incomes of zero from our dataset, because it will interfere with our model accuracy. However, this compromises our model's predicative range: our model will only be able to predict the household income of those who already have jobs (income).

```{r filter-hhincome}
caldata <- caldata %>%
  mutate(hhincome = case_when(
    hhincome == .1 ~ 0,
    TRUE ~ hhincome
  )) #undoing mutation from proposal to return 0 income variables to 0
caldata <- caldata %>%
  filter(hhincome > 5, log(hhincome) < 8)
```


```{r hhincome-distribution}
ggplot(data = caldata, aes(x = hhincome)) +
  geom_histogram(binwidth = 100) +
  labs(title = "Distribution of Household Income", x= "Household Income", y = "Count")
```

Now the distribution of response variable (hhincome) looks like a right skewed normal distribution.

```{r cities, fig.align="center"}
caldata <- caldata %>%
  filter(uscity %in% c("Bakersfield, CA","Bakersfield, CA", "Fresno, CA", "Los Angeles-Long Beach, CA", "Merced, CA","Orange County, CA","Riverside-San Bernardino, CA", "Sacramento, CA", "San Diego, CA", "San Francisco, CA", "San Jose, CA","Santa Barbara-Santa Maria-Lompoc, CA","Santa Cruz-Watsonville, CA","Vallejo-Fairfield-Napa, CA","Ventura, CA"))
ggplot(data = caldata, aes(x = uscity))+ 
  geom_bar() +
  coord_flip()+
  labs(title = "Distribution of Cities in California", x = "City", y = "Count")
```

These immigrants to California arrived to the following cities: Los Angeles-Long Beach, San Francisco, San Diego, Santa Cruz-Watsonville, Bakersfield, Fresno, Merced, Orange County, Riverside-San Bernardino, Sacramento, San Jose, Santa Barbara-Santa Maria-Lompoc, Vallejo-Fairfield-Napa, and Ventura. Given the comparatively small number of cases in which no city was reported, we deleted these instances. This leaves 15 unique locations in California. The majority of immigrants went to LA-Long Beach area. 

#### 2.2.3 Remove Variable "relhead"

It turned out that all values from relhead in our cleaned data were "1" or head. So we will remove this variable, as well as state variables since we are only using California data. We will also remove place data since we are using uscity, and occ since we are using occtype.

```{r remove-relhead}
caldata <- subset(caldata, select = -c(relhead, usstatel, usstate1,usplace1, usplacel,occ) )
```

#### 2.2.4 Mean-center "age" , "usdur1" and "usdurl"

We must center age and usdurl in order to interpret them.

```{r}
mean(caldata$age) #mean age
mean(caldata$usdurl) #mean usdurl
mean(caldata$usdur1) #mean usdur1
```

The mean age in the dataset is 39.43 years ; the mean duration of last US migration is 60.27 months (about 5 years); and the mean duration of first US migration is 43.99 months (less than 3.5 years).

```{r age-usdurl-cent}
caldata <- caldata %>%
  mutate(age=age-mean(age))
caldata <- caldata %>%
  mutate(usdurl=usdurl-mean(usdurl))
caldata <- caldata %>%
  mutate(usdur1=usdur1-mean(usdur1))
```

#### 2.2.5 Remove Obvious Collinear Variable

```{r yrborn-age, fig.align = "center"}
ggplot(data = caldata, mapping = aes(x = yrborn, y = age)) +
  geom_point()+
  labs(title = "Correlation Between Age and Year Born", x = "Year Born", y = "Age")
```

`yrborn` and `age` provide the same information and are perfectly linear, therefore we decided to remove `yrborn` from consideration in the model.


## 3. Multiple Linear Regression Model

In an effort to explain which characteristics of candidates influence their household income, we will be using a multiple linear regression model. Since our response variable is numerical with mulitple potential predictors, this is the best model at our disposal for us to use.

We will consider the potential interaction between principal occupation and number of years of school completed, since those are generally interconnected. We may also consider the interaction between documentation type and occupation type, although the effect may be insignificant.

We will select our model using AIC criteria, because since we're dealing with people, we want to build a model that accounts for volatile human nature and the ever-changing socioeconomic and political climate that could influence someone's household income. AIC is used when we would rather say a variable is a relevant predictor, when in reality it might not be and so in this case, we would rather err on the side of a false positive because we are dealing with a constantly fluctuating issue.

### 3.1 Full Model

```{r model}
full <- lm(hhincome ~ sex + age + statebrn + marstat + edyrs + occtype + usdur1 + usdurl + usdoc1 + uscity, data=caldata)
kable(tidy(full),format="html" ,digits=3)
```

### 3.2 Backward selection

```{r backward-selection}
reduced <- step(full, direction = "backward")
sel_summary <- summary(reduced)
tidy(reduced, exponentiate = FALSE, conf.int = TRUE) %>%
kable(digits = 3, format = "markdown")
```

Using backward selection based on AIC, we narrowed down to 4 variables: sex, edyrs, usdurl and age.


### 3.3 Interactions

To find potential interactions between the 4 variables, we used nested-F test for each of the possible interactions:

```{r int-model}
int_model_1 <- lm(hhincome ~ sex + age + edyrs + usdurl + edyrs*usdurl, data = caldata)
anova(reduced, int_model_1, test = "Chisq") %>% 
  kable(format = "markdown", digits = 3)  # p = 0.653

int_model_2 <- lm(hhincome ~ sex + age + edyrs + usdurl + sex*usdurl, data = caldata)
anova(reduced, int_model_2, test = "Chisq") %>% 
  kable(format = "markdown", digits = 3) # p = 0.034

int_model_3 <- lm(hhincome ~ sex + age + edyrs + usdurl + age*usdurl, data = caldata)
anova(reduced, int_model_3, test = "Chisq") %>% 
  kable(format = "markdown", digits = 3)  # p = 0.201

int_model_4 <- lm(hhincome ~ sex + age + edyrs + usdurl + sex*edyrs, data = caldata)
anova(reduced, int_model_4, test = "Chisq") %>% 
  kable(format = "markdown", digits = 3)  # p = 0.326

int_model_5 <- lm(hhincome ~ sex + age + edyrs + usdurl + age*edyrs, data = caldata)
anova(reduced, int_model_5, test = "Chisq") %>% 
  kable(format = "markdown", digits = 3)  # p = 0

int_model_6 <- lm(hhincome ~ sex + age + edyrs + usdurl + age*sex, data = caldata)
anova(reduced, int_model_6, test = "Chisq") %>% 
  kable(format = "markdown", digits = 3)  # p = 0.376
```

Through nested F-test, we observed significant interactions between age & edyrs and between sex & usdurl, with respective p-values of 0 and 0.034.

### 3.4 Model with Interaction

```{r final-model}
reduced_int <- lm(hhincome ~ sex + age + edyrs + usdurl + age*edyrs + sex*usdurl, data=caldata)
kable(tidy(reduced_int),format="html" ,digits=3)
```

### 3.5 Backward Selection with Interaction

Since we observed 2 pairs of significant interactions, we will do the backward selection again with the new interaction terms.

```{r}
full_int <- lm(hhincome ~ sex + age + statebrn + marstat + edyrs + occtype + usdur1 + usdurl + usdoc1 + uscity + sex*usdurl + age*edyrs , data=caldata)
kable(tidy(full_int),format="html" ,digits=3)
```

```{r}
reduced_int_new <- step(full_int, direction = "backward")
sel_summary <- summary(reduced_int_new)
tidy(reduced_int_new, exponentiate = FALSE, conf.int = TRUE) %>%
kable(digits = 3, format = "markdown")
```

We observed that the variable selection changed. Specifically, usdur1 and usdoc1 are significant, besides the 4 variables and the 2 interaction terms identified previously.





## 4. Check Assumtpions

Before interpreting the model, it is essential to check the assumptions.

### 4.1 Linearity

### 4.1.1 Predicted vs. Factors

```{r scatter-pairs}
pairs(hhincome ~ age + edyrs + usdurl + usdur1 , data = caldata,
      lower.panel = NULL)
```

```{r income-sex-plot}
ggplot(data = caldata, aes(x = as.factor(sex), y = hhincome))+
         geom_boxplot()+
  labs(title = "Boxplot of Household Income against Sex", x = "Sex", y = "Household income")

```

From the above plots, there seems to be a weak linear relationship between the response and predictor variables.

We also observe that there is a distinct linear relationship between usdurl and usdur1, which makes sense, because many people's first migration is their last migration. (they have only migrated to the US once.)

Hence, it is worth looking at VIF of the variables:

#### 4.1.2 Multicollinearity
```{r}
vif(reduced_int_new)
vif(reduced_int)
```

The VIF of usdurl and sex:usdurl are greater than 10, which is alarming. The high VIF of usdurl is likely due to its collinearity with usdur1. However, looking at the VIF for reduced_int model, the VIF for usdurl is still high, meaning that it actually have something to do with the interaction term "sex*usdurl". 

Hence, we should possibly consider dropping either usdurl or sex*usdurl.

#### 4.1.2.1 Drop sex*usdurl
```{r}
full_int_new_1 <- lm(hhincome ~ sex + age + statebrn + marstat + edyrs + occtype + usdur1 + usdurl + usdoc1 + uscity + age*edyrs , data=caldata)
kable(tidy(full_int_new_1),format="html" ,digits=3)

reduced_int_new_1 <- step(full_int_new_1, direction = "backward")
sel_summary <- summary(reduced_int_new_1)
tidy(reduced_int_new_1, exponentiate = FALSE, conf.int = TRUE) %>%
kable(digits = 3, format = "markdown")
```

```{r}
vif(reduced_int_new_1)
```
#### 4.1.2.2 Drop usdurl and Change usdurl*sex to usdur1*sex

```{r}
full_int_new_2 <- lm(hhincome ~ sex + age + statebrn + marstat + edyrs + occtype + usdur1 + usdoc1 + uscity + age*edyrs + sex*usdur1 , data=caldata)
kable(tidy(full_int_new_2),format="html" ,digits=3)

reduced_int_new_2 <- step(full_int_new_2, direction = "backward")
sel_summary <- summary(reduced_int_new_2)
tidy(reduced_int_new_2, exponentiate = FALSE, conf.int = TRUE) %>%
kable(digits = 3, format = "markdown")
```

```{r}
vif(reduced_int_new_2)
```

We observed that neither new models have large VIF, which is good. To choose between the two we need to see if the interaction term sex*usdur1 is significant enough:

```{r}
int_model_7 <- lm(hhincome ~ sex + age + edyrs + usdur1 + age*sex, data = caldata)
anova(int_model_7, reduced_int_new_2, test = "Chisq") %>% 
  kable(format = "markdown", digits = 3)  # p = 0
```

Since the p-value is 0, we have confidence that the interaction sex*usdur1 is significant. Also, we agreed to err on the side of false positive. Hence, we will include more variables, and choose the reduced_int_new_2 model to proceed.



### 4.2 Constant Variance

### 4.2.1 Residuals vs. Factors

```{r add-residuals}
caldata <- caldata %>% 
  mutate(predicted = predict.lm(reduced_int_new_2), residuals = resid(reduced_int_new))
```

Below is the plots of residuals against each quantitative predictor:

```{r factor_residual_scatterplot}
pairs(residuals ~ age + usdur1 + edyrs, data = caldata,
      lower.panel = NULL) 
```

From the pairs plot above, there is a possibility that the constant variance for 'usdur1' is violated, because the scatterplot shows a reserved-fan shape. However, this could also be due to the fact that we have more data points on the lower range as compared to the upper range. Even though, the data looks sparser on the right, the spread is still large.

We believe that this violation of constant variance may result from certain migration patterns- seasonal workers which come to the US for a short period of time would expect to see different household incomes than long-term migrants. We can examine the distribution of duration of first US migration.

```{r usdurl-plot}
ggplot(data = caldata, aes(x = usdur1)) +
  geom_histogram() +
  labs(title = "Distribution of First Migration to US (Mean-Centered)", x = "Duration - 43.99 months (months)")
```

```{r create-seasonal-var}
caldata <- caldata %>%
  mutate(seasonal = case_when(
    usdur1 < 6 - 43.99 ~ "Y", #Workers with duration of migration < 6 months will be considered seasonal
    TRUE ~ "N"
  ))

caldata %>%
  count(seasonal)

ggplot(data = caldata, aes(x = seasonal, y = usdur1)) +
  geom_boxplot() +
  labs(title = "Distribution of Duration of First US Migration", x = "Seasonal Worker?", y = 
         "Duration of First US Migration (Mean-Centered) (months)")
ggplot(data = caldata, aes(x = seasonal, y = hhincome)) +
  geom_boxplot() +
  labs(title = "Distribution of Household Income", x = "Seasonal Worker?", y = 
         "Household Income (dollars)")
```


We create a new categorical variable called `seasonal` to classify migrants with a last US migration shorter than 6 months as seasonal workers. We chose 6 months because anything longer than that should not be considered seasonal, and roughly a quarter of the migrants in the dataset fall into this category.


Hence we fit our new full model:

```{r}
full_seasonal <- lm(hhincome ~ sex + age + statebrn + marstat + edyrs + occtype + usdur1 + usdoc1 + uscity + sex*usdur1 + age*edyrs + seasonal , data=caldata)
kable(tidy(full_seasonal),format="html" ,digits=3)
```

backward selection:

```{r}
reduced_seasonal <- step(full_seasonal, direction = "backward")
sel_summary <- summary(reduced_seasonal)
tidy(reduced_seasonal, exponentiate = FALSE, conf.int = TRUE) %>%
kable(digits = 3, format = "markdown")
```

The backward selection gave the same output as before, indicating that the indicator "seasonal" is not significant enough for us to include it in the final model.

Hence, we proceed with the rest of our assumption checking.


Below is the graph of residuals against categorical predictor:

```{r floorCat_residual}
ggplot(data=caldata,aes(x=as.factor(sex),y=residuals)) + 
  geom_boxplot() + 
  geom_hline(yintercept=0,color="red") +
  labs(x = "Sex", 
       y="Residuals",
       title= "Boxplot of Residuals vs. Sex")

ggplot(data=caldata,aes(x=as.factor(usdoc1),y=residuals)) + 
  geom_boxplot() + 
  geom_hline(yintercept=0,color="red") +
  labs(x = "First US Documentation", 
       y="Residuals",
       title= "Boxplot of Residuals vs. First US Documentation")
```

From the boxplot, the median of each category floats around 0. While the plot of male is relatively symmetrical, that of female is right skewed. There isn't any major concerns

#### 4.2.2 Residual vs. Predicted

```{r residuel-predicted-scatterplot}
caldata <- caldata %>% 
  mutate(predicted = predict.lm(reduced_int_new_2), residuals = resid(reduced_int_new_2))
ggplot(data=caldata,aes(x=predicted, y=residuals)) + 
  geom_point() + 
  geom_hline(yintercept=0,color="red") +
  labs(title="Residuals vs. Predicted Household Income")
```

The residual vs. predicted values scatterplot shows no discernible patterns. There is a clustering of predictions between USD 500 and USD 1000, which makes sense because our input data does not have a lot of rich households.

Overall, the graphs confirm the constant variance assumption is satisfied.


### 4.3 Normality

#### 4.3.1 Histogram of Residuals

```{r histogram-residuals}
ggplot(data=caldata, mapping=aes(x=residuals)) + 
  geom_histogram() +
  labs(title="Distribution of Residuals", x="Residual")
```

#### 4.3.2 Normal-QQ Plot of Residuals

```{r qqplot}
ggplot(data=caldata,mapping=aes(sample=residuals)) +
  stat_qq() +
  stat_qq_line() +
  labs(title="Normal QQ Plot of Residuals")
```

From the histogram, the shape of distribution of residuals has a right skew. The Normal-QQ plot supports this finding, as the left half of the line follows closely the theoretical diagonal, yet the right half deviates from diagonal, showing that the right tail is skewed. Since we have 513 observations, the model is robust enough to tolerate deviation from normal.

### 4.4 Independence

Since we only included observations of heads of household, there will not be dependence caused by similar demographics between family members. However, some interdependence may arise since all observations are geographically close to each other. 

 
## 5. Interpretations

Our final model is 
hhincome = 456.946 + 289.409 sexM - 7.630 age + 24.947 edyrs + 1.122 usdur1 + 166.464 usdoc1Legal resident - 94.505 usdoc1Temporary: Tourist/visitor - 41.715 usdoc1Undocumented +  2.712 age:edyrs - 1.184 sexM:usdur1

Intercept:
Our model expects an average household income of $456.946 for female Mexican migrants who has contract Bracero, at age 39 with no years of education and a duration of 43.99 months for first US migration, all else held constant.

Coefficient:
All else constant, as compared to a female, a male's household income increases by $289.409;
All else constant, with each year of increase in age, the household income decrease by $7.630;
All else constant, with each year of increase in years of education, the household income increases by $24.947;
All else constant, with each month of increase in duration in the US during first migration, the household income increases by $1.122;
All else constant, as compared to contract Bracero, the household income of a legal resident increases by $166.464;
All else constant, as compared to contract Bracero, the household income of a temporary visitor decreases by $94.505;
All else constant, as compared to contract Bracero, the household income of an undocumented person decreases by $41.715.
 

Interaction:

For every increase of one month in duration of first US migration above 43.99 months, we expect to see an increase of around 1.18 dollars in household income for women and a (lesser) increase of 1.12 dollars in income for men.

For every increase of one year of age, we expect to see an increase of around 2.712+24.947= 27.66 dollars in household income for each year of increase in education.

We note that being male increases expected income. This may have to do with the gender wage gap, but more likely has to do with the type of labour immigrants tend to do: often, this is physical labour that men, who on average are larger, are more capable of performing. Similarly, we would expect elderly immigrants to struggle with hard labour, and we note that income falls as age increases. 

As anticipated, more years of education increased income significantly. 

The more time an immigrant had been in the US, the more income they could expect to earn -- this may have to do with the advantages of having networks and stability, as well as overcoming language and cultural barriers. 

## 6. Additional Work