---
title: "Characteristics of Recent Mexican Immigrants to the US"
subtitle: "Regression Analysis"
author: "Ben 10"
date: "November 14, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

```{r load-packages}
library(tidyverse)
library(broom)
library(leaps)
library(knitr)
library(dplyr)
```

```{r}
data <- read_csv("/cloud/project/03-regression-analysis/data.csv")
data
```

#### Qintian

#### Thea

```{r}
caldata <- data %>%
  filter(usstatel == "California")
caldata
```


#### Rachel

#### Lilly

```{r}
caldata %>%
  filter(hhincome < exp(11)) %>%
  ggplot(aes(x = log(hhincome))) +
  geom_histogram()
caldata <- caldata %>%
  mutate(hhincome = case_when(
    hhincome > exp(11) ~ hhincome/300,
    hhincome < exp(11) ~ hhincome
  ))
caldata <- caldata %>%
  filter(hhincome > 5)
ggplot(data = caldata, aes(x = log(hhincome))) +
  geom_histogram()


```

Your regression analysis results go here. At a minimum, the regression analysis should include the following: 

- Statement of the research question and modeling obejctive (prediction, inference, etc.)
- Description of the response variable
- Updated exploratory data analysis, incorporating any feedback from the proposal
- Explanation of the modeling process and why you chose those metohds, incorporating any feedback from the proposal
- Output of the final model
- Discussion of the assumptions for the final model
- Interpretations / interesting findings from the model coefficients
- Additional work of other models or analylsis not included in the final model.

*Use proper headings as needed.*