---
title: "Characteristics of Recent Mexican Immigrants to the US"
subtitle: "Regression Analysis"
author: "Ben 10"
date: "November 14, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

```{r load-packages}
library(tidyverse)
library(broom)
library(leaps)
library(knitr)
library(dplyr)
```

```{r data}
data <- read_csv("/cloud/project/03-regression-analysis/data.csv")
data
```

#### Qintian

#### Thea


```{r}
caldata <- data %>%
  filter(usstatel == "California")
```

```{r caldata}
caldata <- data %>%
  filter(usstatel == "California")
caldata

```


#### Rachel
```{r cities}
caldata <- caldata %>%
  filter(usplace1 %in% c(680, 2840, 4480, 4940, 5945, 6780, 
          6920,6920, 7320,7360,7400,7480,7485,8720,8735))
unique(caldata$usplace1)
```
These immigrants to California arrived to the following cities: Los Angeles-Long Beach, San Francisco, San Diego, Santa Cruz-Watsonville, Bakersfield, Fresno, Merced, Orange County, Riverside-San Bernardino, Sacramento, San Jose, Santa Barbara-Santa Maria-Lompoc, Vallejo-Fairfield-Napa, and Ventura. Given the comparatively small number of cases in which no city was reported, we deleted these instances. This leaves 15 unique locations in California. 


#### Lilly


Your regression analysis results go here. At a minimum, the regression analysis should include the following: 

- Statement of the research question and modeling obejctive (prediction, inference, etc.)
- Description of the response variable
- Updated exploratory data analysis, incorporating any feedback from the proposal
- Explanation of the modeling process and why you chose those metohds, incorporating any feedback from the proposal
- Output of the final model
- Discussion of the assumptions for the final model
- Interpretations / interesting findings from the model coefficients
- Additional work of other models or analylsis not included in the final model.

*Use proper headings as needed.*