---
title: "Characteristics of Recent Mexican Immigrants to the US"
subtitle: "Regression Analysis"
author: "Ben 10"
date: "November 14, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r load-packages}
library(tidyverse)
library(broom)
library(leaps)
library(knitr)
library(dplyr)
```


#### Qintian

#### Thea

```{r}
##caldata <- data %>%
  ##filter(usstatel == "California")
```


#### Rachel

#### Lilly



Your regression analysis results go here. At a minimum, the regression analysis should include the following: 

- Statement of the research question and modeling obejctive (prediction, inference, etc.)
- Description of the response variable
- Updated exploratory data analysis, incorporating any feedback from the proposal
- Explanation of the modeling process and why you chose those metohds, incorporating any feedback from the proposal
- Output of the final model
- Discussion of the assumptions for the final model
- Interpretations / interesting findings from the model coefficients
- Additional work of other models or analylsis not included in the final model.

*Use proper headings as needed.*