---
title: "Characteristics of Immigrants to the US by Type of Immigration Status"
subtitle: "Project Proposal"
author: "Ben 10"
date: "October 27, 2019"
output: github_document
---

```{r}
library(readr)
library(knitr)
library(tidyverse)
library(dplyr)
```

## Section 1. Introduction
In recent years, immigration has been a topic of intense controversy. Given the complexity of immigration law, we are interested in examining patterns in immigration since the beginning of the millenia. The U.S. stance on immigration has grown increasingly strict, not only for "illegal" immigration, but also for legal immigration: naturalization, visa use, lawful permanent residency, refugee status, and asylum seekers. 
For example, the number of refugees accepted to the US fell to 22,491 in 2018 from 84,994 in 2016 according to the Department of State, likely as a result of hard line policies against refugees coming to the US. As immigration policies seem likely to tighten in the lead up to the 2020 elections,###logistic regression on whetehr they got citizenship or not of those who applied... (what prop applied compared to overall. It would be interesting to see what contributed to who applies (Tackett)###

Given the variety of ways to immigrate legally, and the varying processes and timelines for each, we would expect that the characteristics of immigrants using each pathway to differ. By examining the demographics of each immigration pathway, we can determine the trends, if any, that exist in the current immigration infrastructure.

Our research project aims to determine the differences in characteristics between different classes of legal immigration. We hypothesize that there exist statistically significant differences in the populations which recieve different types of visa, and we particularly are interested in region of origin as a way to identify populations that may be targeted by federal policy. 

## Section 2: Exploratory Data Analysis



## Section 2. Analysis plan

```{r}
data <- read_csv(paste0("/cloud/project/02-data/data.csv"))
data
```



## Section 3. Regression Analysis Plan

In an effort to determine which characteristics of candidates influence their approval of a visa, we will be using a logistic regression model. Since our response variable is categorical (i.e., approved or denied), this is the best model at our disposal for us to use.

We will consider the potential interaction between occupation and household income, since those are generally interconnected. 

We will select our model using AIC criteria, because since we're dealing with people, we want to build a model that accounts for volatile human nature and the ever-changing sociopolitical climate that could influence whether or not someone's visa request is approved. AIC is used when we would rather say a variable is a relevant predictor, when in reality it might not be and so in this case, we would rather err on the side of a false positive because we are dealing with a constantly fluctuating issue.

## Section 4. References

Use the glimpse function print a summary of the data frame at the end of your proposal.
Place your data in the /data folder, and add dimensions (number of observations and variables) and the data dictionary (a description of every variable in the dataset) to the README in the folder. Make sure the data dictionary is neatly formatted and easy to read.

Sources:

Statistic used in Introduction:
RPC. (January 31, 2019). Number of refugee admissions in the U.S. from the fiscal year of 1990 to the fiscal year of 2018 [Graph]. In Statista. Retrieved October 29, 2019, from https://www-statista-com.proxy.lib.duke.edu/statistics/200061/number-of-refugees-arriving-in-the-us/

## The Data