---
title: "Characteristics of Immigrants to the US by Type of Immigration Status"
subtitle: "Project Proposal"
author: "Ben 10"
date: "October 27, 2019"
output: github_document
---

```{r}
library(readr)
library(knitr)
library(tidyverse)
library(dplyr)
```

## Section 1. Introduction
In recent years, immigration has been a topic of intense controversy. Given the complexity of immigration law, it is of interest to examine patterns in immigration since the beginning of the millenia. The US stance on immigration has grown increasingly strict, not only for "illegal" immigration, but for legal immigration, such as naturalization, visa use, lawful permanent residency (also referred to as green card use), refugee status, and asylum seekers. For example, the number of refugees accepted to the US fell to 22,491 in 2018 from 84,994 in 2016 according to the Department of State, likely as a result of hard line policies against refugees coming to the US. As immigration policies seem likely to tighten in the lead up to the 2020 elections,###logistic regression on whetehr they got citizenship or not of those who applied... (what prop applied compared to overall. It would be interesting to see what contributed to who applies (Tackett)###

Given the variety of ways to immigrate legally, and the varying processes and timelines for each, we would expect that the characteristics of immigrants using each pathway to differ. By examining the demographics of each immigration pathway, we can determine the trends, if any, that exist in the current immigration infrastructure. Determining, for example, those paths to immigration which are most vital for at risk groups of people, may aid in future analysis of federal administrative policies. 

Our research project aims to determine the differences in characteristics between different classes of legal immigration. We hypothesize that there exist statistically significant differences in the populations which recieve different types of visa, and we particularly are interested in region of origin as a way to identify populations that may be targeted by federal policy. 

## Section 2: Exploratory Data Analysis



## Section 2. Analysis plan

```{r}
data <- read_csv(paste0("/cloud/project/02-data/data.csv"))
data
```

The dataset is from The Mexican Migration Project (MMP). It was created in 1982 by an interdisciplinary team of researchers to further our understanding of the complex process of Mexican migration to the United States. The project is a binational research effort co-directed by Jorge Durand, professor of Social Anthropology at the University of Guadalajara (Mexico), and Douglas S. Massey, professor of Sociology and Public Affairs, with a joint appointment in the Woodrow Wilson School, at Princeton University (US).

Since its inception, the MMP's main focus has been to gather social as well as economic information on Mexican-US migration. The data collected has been compiled in a comprehensive database that is available to the public free of charge for research and educational purposes through this web-site.

The MMP170 Database contains an initial file with general demographic and migratory information for each member of a surveyed household (PERS). Pers170 has 132 variables and 176701 observations, hence is very large. Therefore, we selected 16 variables and filtered out rows that contain N/A's to create a new data table "data". These variables include:

"X1":
"sex" : Sex
  1= male
  2= female
"relhead": Relationship to household head
  1= Head 
  2= Husband/wife 
  3= Son/daughter 
  4= Father/mother 
  5= Brother/sister
  6= Niece/nephew
  7= Uncle/aunt 
  8= Cousin 
  9= Grandfather/grandmother 
  10= Grands on/granddaughter 
  11= Stepson/stepdaughter 
  12= Stepbrother/stepsister 
  13= Stepmother/stepfather 
  14= Son-in-law/daughter-in-law 
  15= Brother-in-law/sister-in-law 
  16= Father-in-law/m other-in-law 
  17= Other blood relative 
  18= Other relative by marriage 
  19= Non-relative 
  20= Adopted/foster child

```{r fix-statebrn}
data <- data %>%
  mutate(statebrn = case_when(
    statebrn == 1 ~ "Aguascalientes", 
    statebrn == 2 ~ "Baja California del Norte", 
    statebrn == 3 ~ "Baja California del Sur",
    statebrn == 4 ~ "Campeche",
    statebrn == 5 ~ "Coahuila",
    statebrn == 6 ~ "Colima",
    statebrn == 7 ~ "Chiapas",
    statebrn == 8 ~ "Chihuahua",
    statebrn == 9 ~ "Mexico City",
    statebrn == 10 ~ "Durango",
    statebrn == 11 ~ "Guanajuato",
    statebrn == 12 ~ "Guerrero",
    statebrn == 13 ~ "Hidalgo",
    statebrn == 14 ~ "Jalisco",
    statebrn == 15 ~ "México",
    statebrn == 16 ~ "Michoacán",
    statebrn == 17 ~ "Morelos",
    statebrn == 18 ~ "Nayarit",
    statebrn == 19 ~ "Nuevo Leon",
    statebrn == 20 ~ "Oaxaca",
    statebrn == 21 ~ "Puebla",
    statebrn == 22 ~ "Querétaro",
    statebrn == 23 ~ "Quintana Roo",
    statebrn == 24 ~ "San Luis Potosí",
    statebrn == 25 ~ "Sinaloa",
    statebrn == 26 ~ "Sonora",
    statebrn == 27 ~ "Tabasco",
    statebrn == 28 ~ "Tamaulipas",
    statebrn == 29 ~ "Tlaxcala",
    statebrn == 30 ~ "Veracruz",
    statebrn == 31 ~ "Yucatán",
    statebrn == 32 ~ "Zacatecas",
    statebrn == 212 ~ "El Salvador",
    statebrn == 9999 ~ "NA")
  )
```

"yrborn": Year of birth
"age": Age
"statebrn": State of birth
1 Aguascalientes
2 Baja California del Norte
3 Baja California del Sur
4 Campeche
5 Coahuila
6 Colima
7 Chiapas
8 Chihuahua
9 Mexico City
10 Durango
11 Guanajuato
12 Guerrero
13 Hidalgo
14 Jalisco
15 México
16 Michoacán
17 Morelos
18 Nayarit
19 Nuevo Leon
20 Oaxaca
21 Puebla
22 Querétaro
23 Quintana Roo
24 San Luis Potosí
25 Sinaloa
26 Sonora
27 Tabasco
28 Tamaulipas
29 Tlaxcala
30 Veracruz
31 Yucatán
32 Zacatecas
212 El Salvador

"marstat": Marital status
  1= Never married 
  2= Married (civil or religious) 
  3= Consensual union 
  4= Widowed 
  5= Divorced 
  6= Separated
"edyrs": School years completed
"occ": Principal occupation

```{r fix-occ}
data %>%
  select(occ) %>%
  distinct() %>%
  arrange(desc(occ))
```
Unemployed/Not in the labor force (10 - 99)
10 Unemployed (seeking work)
20 Homemaker
21 Helps around the house
30 Idle (adult not seeking work and not helping around
the house)
40 School-aged, unspecified
41 School-aged or younger, not in school
42 Student
43 Student and worker
50 Retired, unspecified
51 Retired w/o pension
52 Retired with pension
53 Disability retirement
54 Pensioner who works
60 Other, unspecified (disabled, incarcerated, tourist and other)
61 Disabled, ill
62 Incarcerated
63 Tourist
64 On welfare
99 Other, not in workforce
Professionals (110 - 119)
110 Architects; civil, chemical, industrial engineers; etc.
111 Physicists; astronomers; mathematicians;
statisticians; actuaries
112 Chemists and pharmacists
113 Physicians; dentists; optometrists; nutritionists; professional nurses, etc.
114 Biologists; ecologists; marine biologists; etc.
115 Agriculturalists; veterinarians; and professionals in
forestry and fisheries
116 Social scientists, lawyers, and psychologists
117 Economists; business administrators; CPAs; etc.
118 Religious professionals
119 Other professionals
Technical workers (120 - 129)
120 Draftsmen; equipment technicians; video and sound technicians; etc.
121 Technicians in physics, mathematics, statistics and actuarial science
122 Medical technicians: nurse=s aides, dental technicians
123 Lab technicians (chemical, biological, pharmacological, and ecological)
124 Technicians in agriculture, veterinary sciences, forestry, fisheries, etc.
125 Technicians in the social sciences, accounting, administration and tourism
126 Technicians in religious activities
129 Other technicians
Educators (130 - 139)
130 Professors in universities and other institutions of higher learning
131 Professors/teachers in high school or the equivalent
MMP Codebook (1-170)
132 Professors/teachers in junior high school or the equivalent
133 Professors/teachers in grammar school or the equivalent
134 Professors/teachers in preschool
135 Professors/teachers in special education
136 Professors and instructors in arts, administration,
vocational arts, technical education and sports 139 Other educational workers
Occupations in the arts, performances and sports
(140 - 149)
140 Writers; critics; journalists; editors; etc.
141 Composers; singers; musicians; actors; dancers;
etc.
142 Painters; sculptors; illustrators (fine artists);
designers; choreographers; etc.
143 Directors; producers; broadcasters; etc.
144 Athletes
145 Sports referees, umpires and coaches
146 Cartoonists; magicians; clowns; etc.
149 Other artists
Administrators and directors in both public and private sector (210 - 219)
210 Government administrators and legislators 211 Presidents, directors, senior managers, large
factory owners
212 Specialized directors, managers and administrators
(for owners refer to 214, 215, and 710).
213 Directors of political, union and civil organizations
(non-profit).
214* Small and medium-sized factory owners (for
retail establishments see 710; for service
establishments, see 215)
215* Owners of small and medium-sized service
establishments (for retail establishments see 710; for
factories, see 214)
219 Other administrators such as entrepreneurs,
managers, and directors, when no further specification was provided
Agriculture, husbandry, forestry/fisheries workers
(410 - 419)
410 Agricultural workers
411 Husbandry workers
412 Workers in both agriculture and husbandry
413 Forestry workers
414 Hunters; games men; trappers; etc.
415 Fishery or marine workers
416 Workers in activities associated with agricultural or
marine products
417 Foremen, overseers and other control persons of
agricultural, husbandry or fishery activities
419 Other agriculture, husbandry, forestry, fishery
workers
- D1 -
February 2019
Manufacturing /repair supervisors (510 - 519)
510 Food, beverage and tobacco production supervisors
511 Mine, quarry and well supervisors
512 Textile and leather production supervisors
513 Wood and paper production or printing supervisors
514 Electrical, electronic, or metallurgical production
supervisors
515 Ceramic, tile, glass or other mineral production
supervisors
516 Construction, installation, maintenance and finishing
supervisors
517 Electrical generation, installation, repair and
maintenance supervisors (including
telecommunications equipment)
518 Chemical, petroleum, oil, and plastics production
supervisors
519 Other supervisors including those in unspecified
industry
Manufacturing /repair skilled workers (520 - 529). For helpers, aides, apprentices, and trainees see 540- 549.
520 Food, beverage and tobacco production workers, including cooks in establishments.
521 Mine, quarry and well workers
522 Textile and leather production workers. (Examples:
tailors, upholsterers, cobblers, embroiderers, lithographers, seamstresses; for unskilled finishing work, see 542; for clothing designers, see 142.)
523 Wood and paper production or printing workers. (Examples: carpenter, cabinetmaker, linotypist, film developer, other skilled carpentry work)
524 Metal production and treatment workers; vehicle, machinery and equipment repair. (Examples: casters, lathe operators, boilermakers, welders, jewelers, goldsmiths, locksmiths, metal polishers, tool sharpeners, blacksmiths, metal forgers, refrigerator repair people, musical instrument repair people)
525 Ceramic, tile, glass or other mineral production workers. (Examples: potters, glasscutters)
526 Construction, installation, maintenance and finishing workers. (Examples: bricklayers, house painters, plasterers, roofers, floor polishers, plumbers, parts installers)
527 Electrical equipment, electronics and telecommunications installation and repair workers. (Examples: electricians, television/radio repair people).
528 Chemical, petroleum, oil, and plastics production workers
529 Other craftsmen or manufacturing workers, including those in unspecified industry
Manufacturing/repair heavy equipment operators
(530 - 539)
530 Food, beverage and tobacco production equipment operators
531 Mine, quarry and well equipment operators
532 Textile and leather production equipment operators
533 Wood and paper production or printing equipment
operators (includes furniture production).
534 Metallurgical or automotive production or repair
equipment operators. (Examples: assembling
machine operators, rollers, fitters)
535 Ceramic, tile, glass or other mineral production
equipment operators
536 Construction equipment operators
537 Energy, pump or refrigeration equipment operators
538 Chemical, petroleum, oil, and plastics production
equipment operators
539 Other operators of heavy machinery and equipment,
including those in unspecified industry
Manufacturing/repair unskilled workers (540 - 549).
540 Food, beverage and tobacco production unskilled workers
541 Mine, quarry and well unskilled workers
542 Textile and leather production unskilled workers.
(Includes garment finishing work, e.g. sewing
buttons.)
543 Wood and paper production or printing unskilled
workers (includes furniture production)
544 Metallurgical or automotive production or repair
unskilled workers
545 Ceramic, tile, glass or other mineral production
unskilled workers
546 Construction unskilled workers
547 Electrical equipment, electronics and
telecommunications installation and repair unskilled
workers
548 Chemical, petroleum, oil, and plastics production
unskilled workers
549 Other unskilled workers including those in
unspecified industry (includes unspecified helpers or trainees)
Transportation workers (550 - 559)
550 Industrial vehicle operators/drivers. (Examples: crane operators, tractor drivers, reapers, lawn mowers)
551 Railroad conductors and workers
552 Truck drivers and land-transport drivers (see also
712) and passenger vehicle drivers 553 Air-transport pilots
554 Maritime captains, pilots and workers 555 Operator of animal driven cart
559 Other conductors, drivers, pilots
Service and administration supervisors (610 - 619). Includes department chiefs, coordinators, and supervisors.
610 Health, social services, education and justice services supervisors
611 Accounting, finance, human resources, library services supervisors
612 Communications and transportation services supervisors
613 Statistics, information, publicity and research services supervisors
614 Public administration supervisors and infrastructure supervisors
615 Culture and recreation services supervisors
616 Restaurant, store, and hotel services supervisors 617 Agriculture, forestry and fisheries service
supervisors
MMP Codebook (1-170)
- D2 -
February 2019

618 Other department supervisors
619 Other workers who perform similar activities,
including those in unspecified industry
Administrative and support workers (620 - 629)
620 Secretaries; typists; data entry, recorders; etc.
621 Cashiers; collectors; ticket sellers; etc.
622 Record-keepers for stores and warehouses
623 Receptionists; travel agent; interviewers; etc.
624 Telephone and telegraph operators
625 Postal and messenger workers
626 Dispatchers; transportation coordinators
627 Other administrative service workers who perform
routinary or simple tasks.
629 Other related workers, including generic office
workers and public servants when no further specification was provided
Sales workers (710 - 719)
710 Merchants in retail establishments, retail business owners and owners of small businesses. (For factory owners, see 214; for owners of medium and large service establishments, see 215).
711 Workers in retail establishments. (Examples: clerks, dispatchers)
712 Distributors or demonstrators in retail establishments, including delivery workers who may or may not also be drivers (see also 552)
713 Sales agents or representatives; brokers; insurance and real estate agents; auctioneers; etc.
719 Other retail workers, including sales people (unknown whether or not person works in an establishment).
Ambulatory workers (720 - 729) (those who work in their own house are included in the previous group)
720 Ambulatory salespeople: toys, lottery tickets, household goods, paper, other inedible items
721 Ambulatory service workers: food vendors, shoe shiners, car/windshield washers, street performers
729 Other ambulatory workers, self-employed day laborers
Personal services workers in establishments; (not in private households) (810 - 819)
810 Innkeepers; bartenders; waiters; flight attendants.
811 Launderers; pressers; and other clothes-cleaning
service workers
812 Doormen; concierges; elevator operators; bellboys;
cleaning workers; gardeners; movers; dishwashers
813 Barbers; hair stylists; etc.
814 Workers in car rental, and other movable rental establishments
815 Party planners; tour guides; event organizers; caregivers in institutions (except nurses, see 113 and 122).
816 Morticians; funeral home workers; etc.
819 Other personal service worker: e.g., parking lot
attendants
Domestic services workers (820)
820 Domestic services workers; caregivers, drivers,
gardeners, butler, and other service workers in private households, e.g. baby sitter.
Protection services workers (830 - 839)
830 Security personnel; police officers; watchmen, firefighters.
831 Armed forces personnel
839 Other related workers
Unspecified/unknown occupations
MMP Codebook (1-170)
- D3 -
February 2019
9999
Other unspecified occupation; unknown

"hhincome" :
"usstate1": First US mig: State of residence
"usstatel": Latest US mig: State of residence
"usplace1": First US mig: City of residence
"usplacel": Latest US mig: City of residence
"usdur1": First US mig: Duration (in months)
"usdurl": Latest US mig: Duration (in months)

## Section 3. Regression Analysis Plan

## Section 4. References

Use the glimpse function print a summary of the data frame at the end of your proposal.
Place your data in the /data folder, and add dimensions (number of observations and variables) and the data dictionary (a description of every variable in the dataset) to the README in the folder. Make sure the data dictionary is neatly formatted and easy to read.

Sources:

Statistic used in Introduction:
RPC. (January 31, 2019). Number of refugee admissions in the U.S. from the fiscal year of 1990 to the fiscal year of 2018 [Graph]. In Statista. Retrieved October 29, 2019, from https://www-statista-com.proxy.lib.duke.edu/statistics/200061/number-of-refugees-arriving-in-the-us/

## The Data